### 2017-12-10 paper/jfp-submission/response.txt

Indeed, we discuss the limitations of random generation in section
5.8. We have added a sentence there describing how we could replace
random generation with constraint-based generation.

Reviewer 2
> 5. your approximation of functions to fun is rather strong, and it is
> unclear to me how well that works. I guess since all executions start
> from the beginning of the programs keeps everything rather concrete,
> so that you do not need to invent functions. But I am not sure, and I
> would like to see a thorough discussion. I am fine with novice
> programs being short and simple since that is quite likely to be the
> case when they resort to Nanomaly, but when you do that, it should be
> clear that you do.  What may be related to this is the idea to deal
> with Currying. What about functions like foldr of which the result is
> a type variable? I know that some static work on error diagnosis
> sometimes runs into a problem of some sort and since you work at the
> dynamic level this may not be the case for you. But do say and explain
> so.

> 2. what is the limit to the size of programs you can reasonably
> handle?  What if we have a large program, is it possible to have a
> different starting point? So does a trace have to start at the start
> of the program? I guess this is important since if you start with a
> higher order function then you need to dynamically invent a function,
> and then your approximation might well be in the way. This is
> something I am really curious about.  Now you have 5% Non-parametric
> function types. Does that change in this situation?

We discuss the approximation of functions to `fun` in the
"Non-Parametric Function Type" subsection of 5.3, it is the other major
limitation of our work (the first being overloading). We do in fact have
to invent functions sometimes (the definition of `gen` says that these
must be constant functions), and this interacts very poorly with
higher-order functions like `foldr`. We have added another example to
the discussion in 5.3 that illustrates the issue.

### 2017-12-10 paper/jfp-submission/safe.tex
more response

\paragraph{Ad-Hoc Polymorphism}
%
We found that for 5\% of all programs \toolname got stuck when it tried to
compare two holes.
%
\ocaml provides polymorphic equality and comparison operators,
overloading them for each type.
%
While convenient to use, they pose a challenge for \toolname's
combination of execution and inference.
%
For example, consider the \hbox{@n <= 0@} test in our @fac@ example.
%
The @<=@ operator is polymorphic, but in this case we can make progress because
the literal @0@ is not.
%
Suppose, however, we parameterized @fac@ by a lower bound, \eg
%
\begin{code}
  let rec fac n m =
    if n <= m then
      true
    else
      n * fac (n - 1) m
\end{code}
%
When given @fac@, \toolname will generate two fresh holes
$\nu_1[\alpha_1]$ and $\nu_2[\alpha_2]$ and proceed directly into the
@n < m@ comparison.
%
We cannot (yet) instantiate either hole because we have no constraints
on the $\alpha$s (we know they must be equal, but nothing else), and
furthermore we do not know what constraints we may encounter later on in
the program.
%
Thus, we cannot perform the comparison and proceed, and must give up our
search for a witness, even though one obviously exists, any pair of |n|
and |m| such that |n <= m| is false.

@@

@@

\paragraph{Summary}
%
Our investigation suggests that the vast majority of programs
for which we fail to find a witness do not, in fact, admit a
witness.
%
% Overall, it appears that our random instantiation of holes is well-suited
% to finding type errors in student programs.
These programs were generally cases where \ocaml's type system was
overly conservative.
%
Of course, the conservatism is somewhat justified as each case pointed
to code that would be difficult to use or maintain;
%
it would be interesting to investigate how demonstrate these issues in an
intuitive manner.

>>>>

\paragraph{Ad-hoc Polymorphism}
%
We found that for 5\% of all programs \toolname got stuck when it tried to
compare two holes.
%
\ocaml provides polymorphic equality and comparison operators,
overloading them for each type.
%
While convenient to use, they pose a challenge for \toolname's
combination of execution and inference.
%
For example, consider the \hbox{@n <= 0@} test in our @fac@ example.
%
The @<=@ operator is polymorphic, but in this case we can make progress because
the literal @0@ is not.
%
Suppose, however, we parameterized @fac@ by a lower bound, \eg
%
\begin{code}
  let rec fac n m =
    if n <= m then
      true
    else
      n * fac (n - 1) m
\end{code}
%
When given @fac@, \toolname will generate two fresh holes
$\nu_1[\alpha_1]$ and $\nu_2[\alpha_2]$ and proceed directly into the
@n < m@ comparison.
%
We cannot (yet) instantiate either hole because we have no constraints
on the $\alpha$s (we know they must be equal, but nothing else), and
furthermore we do not know what constraints we may encounter later on in
the program.
%
Thus, we cannot perform the comparison and proceed, and must give up our
search for a witness, even though one obviously exists, any pair of |n|
and |m| such that |n <= m| is false.

@@

The $\tfun$ abstraction is also problematic when we have to generate new
functions. Consider the following @pipe@ function that composes a list
of functions.
%
\begin{code}
  let pipe fs =
    let f a x = a x in
    let base a = a in
    List.fold_left f base fs
\end{code}
%
In the folding function @f@, the student has applied the accumulator @a@
to the new function @x@, rather than composing the two.
%
The \ocaml compiler again detects a cyclic typing constraint and rejects
the program, but \toolname is unable to produce a witness.
%
In this case the issue lies in the fact that the safety of a call to
@pipe@ is determined by its arguments: the call @pipe [(fun x -> 1)]@ is
safe, but the call @pipe [(fun x -> 1); (fun x -> 2); (fun x -> 3)]@
will get stuck when we try to reduce @1 (fun x -> 3)@.
%
Unfortunately, \toolname is unable to synthesize such a witness because
of the abstraction to $\tfun$.
%
Specifically, this abstraction forces our hand inside $\gensym$; we do
not know the what the input and output types of the function should be,
so the only safe thing to do is to generate a function that accepts any
input and returns a value of a yet-to-be-determined type.
%
Thus, \toolname's lenient instantiation of holes prevents it from
discovering a witness for @pipe@.

@@

\paragraph{Summary}
%
Our investigation suggests that the majority of programs
for which we fail to find a witness do not, in fact, admit a
witness under \toolname's semantics.
%
% Overall, it appears that our random instantiation of holes is well-suited
% to finding type errors in student programs.
These programs were generally cases where \ocaml's type system was
overly conservative.
%
Of course, the conservatism is somewhat justified as each case pointed
to code that would be difficult to use or maintain;
%
it would be interesting to investigate how demonstrate these issues in an
intuitive manner.
